Local PCA Shows How the Effect of Population Structure Differs Along the Genome

### 2016-08-13 writeup/local_pca_paper.tex
plots for drosophila recomb and mds

As depicted in Figure \ref{fig:diagram}, the general steps to the method are:
(1) divide the genome into windows,
(2) summarize the patterns of relatedness in each window,
(3) measure dissimilarity in relatedness between each pair of windows,
(4) visualize the resulting dissimilarity matrix,
and (5) combine similar windows to more accurately visualize local effects of population structure.
Details of how we carried these out are given in the Methods.

@@

We applied the method to genomic datasets with good geographic sampling
of African populations of the fruit fly \textit{Drosophila melanogaster},
a worldwide dataset of humans,
and a range-wide dataset of the partially selfing weedy annual plant \textit{Medicago truncatula}.
In these three species, PCA plots vary along the genome in a systematic way, showing strong chromosome-scale correlations.
This implies that variation is due to meaningful variation in the effects of population structure, 
since noise due to randomness in choice of local genealogical trees
are not expected to show long distance correlations. 
Below, we discuss the results and likely underlying causes.

>>>>

As depicted in Figure \ref{fig:diagram}, the general steps to the method are:
(1) divide the genome into windows,
(2) summarize the patterns of relatedness in each window,
(3) measure dissimilarity in relatedness between each pair of windows,
(4) visualize the resulting dissimilarity matrix using multidimensional scaling (MDS),
and (5) combine similar windows to more accurately visualize local effects of population structure using PCA.
Details of how we carried out each step are given in the Methods,
and code to run the analyses is provided as an R package at \url{https://github.com/petrelharp/local_pca}.

@@

We applied the method to genomic datasets with good geographic sampling:
African populations of the fruit fly \textit{Drosophila melanogaster},
a worldwide dataset of humans,
and a range-wide dataset of the partially selfing weedy annual plant \textit{Medicago truncatula}.
In all three species, PCA plots vary along the genome in a systematic way, showing strong chromosome-scale correlations.
This implies that variation is due to meaningful heterogeneity in a biological process,
since noise due to randomness in choice of local genealogical trees
are not expected to show long distance correlations. 
Below, we discuss the results and likely underlying causes.
